
---
title: Cluster Strings 
author: Christian Saal
date: '`r Sys.Date()`'
slug: []
categories: []
tags: [R]
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300)
knitr::opts_knit$set(global.par = TRUE)
```


# Distance between strings

> Edit distance, also called Levensthein distance, is a measure of the number of primary edits that
> would need to be made to transform one string into another.
> -- <cite>[Jack Davies][1]</cite>

[1]: https://www.r-bloggers.com/2019/04/natural-language-processing-in-r-edit-distance/

Here is a small example where the function `adist()` from the R Utils package that is part of the R
Distribution. 

1. we create a vector containing six words
2. next we calculate the edit distance
3. then using a hierachical cluster analysis to find similarities

```{r}

s_vec   <- c("alcool","alcohol","alcoholic","brandy","brandie","cachaça")
dist    <- adist(s_vec)
hc      <- hclust(as.dist(dist))
dfClust <- data.frame(s_vec, cutree(hc, k=3))

dfClust

```

# Give it a try with soccer data

The data is extracted from eventdata. I subset all episodes or sequences that ends with a SHOT or a
GOAL and sample a small amount of them. Now i can use these attacking sequences to run a cluster analysis.

```{r, include=FALSE}

setwd("~/Desktop/Shortsave/webseite/content/post/2021-05-07-levenshtein/")
load("data/sample.Rdata")
library(stringdist)

```

First we inspect the data in a short way.

1. show me the first six vectors of strings
2. there are 40 sequences within the dataset 
3. 22 of them are unique, that means there are duplicates in there

```{r}

# inspect data
head(mydat)
length(mydat)
length(unique(mydat))

```

Next i create a distance matrix. Here i used the package stringdist. In this package we have the
option to choose several distance metrics as describes in `help("stringdist-metrics")`. Here i choose
the Jaro-Winkler distance. 

```{r}

dist           <- stringdistmatrix(mydat, method="jw")
hc             <- hclust(as.dist(dist))
dfClust        <- data.frame(mydat, cutree(hc, k=2))
names(dfClust) <- c('event_type','cluster')

```

Displaying the results shows that the first cluster contains very short sequences. There are some
DUEL events in there and two sequences ending by a goal. The second cluster includes sequences that
end with a SHOT  and one with a GOAL and are longer sequences.  

```{r}

cluster1 <- dfClust[which(dfClust$cluster == 1),]
cluster2 <- dfClust[which(dfClust$cluster == 2),]

counts1 <- sort(table(cluster1$event_type))
counts2 <- sort(table(cluster2$event_type))

kable(counts1)
kable(counts2)

```

# Summary

Here i just play around with the possibility of cluster strings. According to the event data it
could be useful to do that with attacking sequences of a whole season and then analyse the sequences
by club. 
